---
title: "Aging Dependence Data Cleaning"
author: "Analeidi Barrera and Colleen Minnihan"
date: "2/16/2021"
output: html_document
---


```{r}
#Path to the aging data
path_data = '/Users/ColleenMinnihan/Desktop/Survival_Analysis/ICPSR_08719/DS0001/08719-0001-Data.por'
```

```{r}
library(haven)
library(dplyr)
data <- read_por(path_data)
```


#==================================


```{r}
library(survminer)
library(survival)
```


```{r}
library(haven)
library(dplyr)
data <- read_por('/Users/analeidibarrera/Desktop/ProjectInformation/ICPSR_08719/DS0001/08719-0001-Data.por')
```


```{r}
#selecting covariates

aging <- data %>% select(TLKCHL2, TLKCHL3, TLKCHL4, SEECHL2, SEECHL3, SEECHL4, MLCHL2R, MLCHL3R, MLCHL4R, CHLTME2R, CHLTME3R, CHLTME4R, WKSWORK2, WKSWORK3, WKSWORK4, MARSTAT2, MARSTAT3, MARSTAT4, DIFDRES2,DIFDRES3, DIFDRES4, BATH2R, BATH3R, BATH4R, TOILET2R, TOILET3R, TOILET4R, EAT2R, EAT3R, EAT4R, DIFPREP2, DIFPREP3, DIFPREP4, GET2R, GET3R, GET4R, DIFLTHW2, DIFLTHW3,  DIFLTHW4, DIFHSWK2, DIFHSWK3 , DIFHSWK4 , DIFTEL2 , DIFTEL3 , DIFTEL4,DIFMNY2, DIFMNY3, DIFMNY4, DIFSHOP2, DIFSHOP3, DIFSHOP4, DIFOUT2, DIFOUT3, DIFOUT4, DIFWALK2, DIFWALK3, DIFWALK4, BIRTHYR, RACER, SEX,INCFAMR,EDUC)

#creating a "how quickly can children get there" variable with 4 levels: < 1 hr, >= 1 hr, no children, children in household
aging <- aging %>%
  mutate(children_dist = ifelse(CHLTME2R == '1' | CHLTME2R == '2' | CHLTME2R == '3', 'less than 1 hr', ifelse(CHLTME2R == '4' | CHLTME2R == '5' | CHLTME2R == '6', 'at least 1 hr',ifelse(CHLTME2R == '7','no children',ifelse(CHLTME2R == '8', 'children in house', NA))))) %>%
  mutate(see_children = ifelse(SEECHL2 == '000'|SEECHL2 == '011'|SEECHL2 == '012'|SEECHL2 == '021'|SEECHL2 == '022'|SEECHL2 == '031'|SEECHL2 == '032'|SEECHL2 == '041'|SEECHL2 == '051'|SEECHL2 == '061'|SEECHL2 == '071'|SEECHL2 == '081'|SEECHL2 == '091'|SEECHL2 == '101'|SEECHL2 == '111'|SEECHL2 == '121'|SEECHL2 == '131',"less than 1 month",ifelse(SEECHL2 == '013'|SEECHL2 == '014' |SEECHL2 == '023' |SEECHL2 == '033'|SEECHL2 == '042'|SEECHL2 == '043'|SEECHL2 == '052'|SEECHL2 == '053'|SEECHL2 == '062'|SEECHL2 == '063'|SEECHL2 == '072'|SEECHL2 == '073'|SEECHL2 == '082'|SEECHL2 == '083'|SEECHL2 == '092'|SEECHL2 == '093'|SEECHL2 == '102'|SEECHL2 == '103'|SEECHL2 == '112'|SEECHL2 == '113'|SEECHL2 == '122'|SEECHL2 =='123',"1 month to 1 year", ifelse(SEECHL2 == '024'|SEECHL2 == '034'|SEECHL2 == '044'|SEECHL2 == '054'|SEECHL2 == '064'|SEECHL2 == '074'|SEECHL2 == '084'|SEECHL2 == '094'|SEECHL2 == '104'|SEECHL2 == '114'|SEECHL2 == '124',"greater than 1 year",NA)))) %>%
  mutate(fam_income84 = ifelse(INCFAMR=='0'|INCFAMR=='1'|INCFAMR=='2','less than 10k',ifelse(INCFAMR=='3'|INCFAMR=='4','10k to 19k',ifelse(INCFAMR=='5'|INCFAMR=='6'|INCFAMR=='7'| INCFAMR=='8','20k+',NA))))

  
  
   # mutate(see_children = ifelse(SEECHL2 == c('000','011','012','021','022','031','032','041','051','061','071','081','091','101','111','121','131'),"less than 1 month",ifelse(SEECHL2 == c('013','014','023','033','042','043','052','053','062','063','072','073','082','083','092','093','102','103','112','113','122','123'),"1 month to 1 year", ifelse(SEECHL2 == c('024','034','044','054','064','074','084','094','104','114','124'),"greater than 1 year",NA))))

  #mutate(temp_SEECHL2 = as.numeric(strsplit(as.character(SEECHL2), "")[[1]]))
 # mutate(see_children = ifelse([3] == 1 &)
         

#less than 1 month
#1 month to a year
#greater than a year

as.numeric(strsplit(as.character(aging$SEECHL2), "")[1])

aging$SEECHL2[1]
```

```{r}
n=c(3,2,1)

n[1]


a <- 6547
a <- as.numeric(strsplit(as.character(a), "")[[1]])

a[2]
```




### Creating a KM curve for bathing (BATH2R), using the toilet(TOILET2R), and eating(EAT2R)

In our data we have the following categories:

1: Independent

2: Dependent

3: Don't know,... refused, does not do for other than health reasons


*For the purpose of this analysis, we are assuming that a (3) or no data means that nothing has changed from the previous information obtained. If no previous information was recorded, we assume the person is independent. 

**The edge case of (1, _, 2) is a left censoring scenario since we only have an upper bound. We are assuming the above assumption is true

The participants that we are considering are the following:
- They must have started as independent at whatever year we had concrete information (2 or 1) on them
 

#### Creating a subset data with the targeted dependence/response variables

```{r}
#BATH
bath <- aging %>% filter(BATH2R != "2" & (BATH2R == "1" |BATH3R == "1" | BATH4R == "1") & !(BATH3R == "2" & BATH4R == "1"))

#TOILET
toilet <- aging %>% filter(TOILET2R != "2" & (TOILET2R == "1" |TOILET3R == "1" | TOILET4R == "1") & !(TOILET3R == "2" & TOILET4R == "1"))

#EAT 
eat <- aging %>% filter(EAT2R != "2" & (EAT2R == "1" |EAT3R == "1" | EAT4R == "1") & !(EAT3R == "2" & EAT4R == "1"))

```


The following are the data sizes that we are working with:

Bathing or showering dependence = 5207 participants

```{r}
count(bath)
```

Using or getting to the toilet dependence = 5619 participants

```{r}
count(toilet)
```


Eating dependence = 5904 participants
```{r}
count(eat)
```

####Creating the Status variable

1: Exact data 
0: Censored data



```{r}

#Bath
bath$Status <- ifelse(bath$BATH3R=="2" |bath$BATH4R=="2", 1, 0)

#Toilet
toilet$Status <- ifelse(toilet$TOILET3R=="2" |toilet$TOILET4R=="2", 1, 0)

#Eat
eat$Status <- ifelse(eat$EAT3R=="2" |eat$EAT4R=="2", 1, 0)
```



#### Creating an age variable depicting the age of dependence or censoring


```{r}
#bath data
bath$age <- ifelse(bath$BATH4R== "1", 1990-as.integer(as.character(bath$BIRTHYR)), ifelse(bath$BATH3R== "1", 1988 - as.integer(as.character(bath$BIRTHYR)), 1986 -as.integer(as.character(bath$BIRTHYR))))

bath$age <- ifelse(bath$BATH3R== "2", 1988- as.integer(as.character(bath$BIRTHYR)), ifelse(bath$BATH4R== "2", 1990-as.integer(as.character(bath$BIRTHYR)), bath$age))

#toilet data
toilet$age <- ifelse(toilet$TOILET4R== "1", 1990-as.integer(as.character(toilet$BIRTHYR)), ifelse(toilet$TOILET3R== "1", 1988 - as.integer(as.character(toilet$BIRTHYR)), 1986 -as.integer(as.character(toilet$BIRTHYR))))

toilet$age <- ifelse(toilet$TOILET3R== "2", 1988- as.integer(as.character(toilet$BIRTHYR)), ifelse(toilet$TOILET4R== "2", 1990-as.integer(as.character(toilet$BIRTHYR)), toilet$age))

#eating data
eat$age <- ifelse(eat$EAT4R== "1", 1990-as.integer(as.character(eat$BIRTHYR)), ifelse(eat$EAT3R== "1", 1988 - as.integer(as.character(eat$BIRTHYR)), 1986 -as.integer(as.character(eat$BIRTHYR))))

eat$age <- ifelse(eat$EAT3R== "2", 1988- as.integer(as.character(eat$BIRTHYR)), ifelse(eat$EAT4R== "2", 1990-as.integer(as.character(eat$BIRTHYR)), eat$age))
```

Bath:
```{r}
#Filtering out the participants that did not give their birth year
bath <- bath %>% filter(age > 0)
```

Toilet:
```{r}
#Filtering out the participants that did not give their birth year
toilet <- toilet %>% filter(age > 0)
```

Eat:
```{r}
#Filtering out the participants that did not give their birth year
eat <- eat %>% filter(age > 0)
```



#### Generating the Kaplan-Meier Curve for bathing dependence

```{r}
KMbath = survfit(Surv(age, Status)~1, data = bath)
KMbath
plot(KMbath, conf.int=FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1))
```


####Parametric models for bathing dependence

Defining our parametric models to see which one fits best
```{r}
weibullBath = survreg(Surv(age, Status)~1, dist = "weibull", data = bath)
expBath = survreg(Surv(age, Status)~1, dist = "exponential", data = bath)
logNormalBath = survreg(Surv(age, Status)~1, dist = "lognormal", data = bath)
normalBath = survreg(Surv(age, Status)~1, dist = "gaussian", data = bath)
```


Creating a plot for the following models

Bathing Weibull:

```{r}
curve(1-pweibull(x, shape = 1/0.0499, scale = exp(4.55815)), xlim= c(min(bath$age), max(bath$age)), ylab="Survival", xlab = "Age", col = "purple")
```



Bathing Exponential ?????:
```{r}
summary(expBath)

curve(1-pexp(x, rate = 1/exp(6.7426)), xlim= c(min(bath$age), max(bath$age)), ylab="Survival", xlab = "Age", col = "orange")
```

Bathing Normal:
```{r}
summary(normalBath)
curve(1-pnorm(x, mean = 93.7955, sd = 7.07), xlim= c(min(bath$age), max(bath$age)), ylab="Survival", xlab = "Age", col = "red")
```

Bathing Log-Normal:
```{r}
summary(logNormalBath)

curve(1-plnorm(x, meanlog = 4.5439, sdlog = 0.083), xlim= c(min(bath$age), max(bath$age)), ylab="Survival", xlab = "Age", col = "blue")
```

Now compare all of the parametric models to the KM curve to see which is visually the best fit:
```{r}
plot(KMbath, conf.int=FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1))

curve(1-pweibull(x, shape = 1/0.0499, scale = exp(4.55815)), col = "purple", add = TRUE)

curve(1-pexp(x, rate = 1/exp(6.7426)), col = "orange", add = TRUE)

curve(1-pnorm(x, mean = 93.7955, sd = 7.07), col = "red", add = TRUE)

curve(1-plnorm(x, meanlog = 4.5439, sdlog = 0.083), col = "blue", add = TRUE)
```

From the looks of the plot, it appears that the normal and log-normal curves provide the best estimate of the KM curve.  Choosing between normal and log-normal is too close to tell, so we may need to use AIC.

AIC to test which parametric model fits our bath data best:
```{r}
#AIC = 2(P-L)

#AIC for Weibull
#summary(weibullBath)
2*(2+2445.5)

#AIC for Exponential
#summary(expBath)
2*(1+3879.1)

#AIC for Normal
#summary(normalBath)
2*(2+2417.5)

#AIC for log-normal
#summary(logNormalBath)
2*(2+2414.1)
```

The AICs are quite similar for both normal and log-normal models, meaning they provide a similar fit, but the log-normal model has a slightly smaller AIC, so we would go with that model for our data.

#### Generating the Kaplan-Meier Curve for toileting dependence

```{r}
KMtoilet = survfit(Surv(age, Status)~1, data = toilet)
KMtoilet
plot(KMtoilet, conf.int=FALSE, ylab="Survival" , xlab="Age", xlim = c(min(toilet$age), max(toilet$age)), ylim = c(0,1))
```

####Parametric models for toileting dependence

Defining our parametric models to see which one fits best
```{r}
weibullToilet = survreg(Surv(age, Status)~1, dist = "weibull", data = toilet)
expToilet = survreg(Surv(age, Status)~1, dist = "exponential", data = toilet)
logNormalToilet = survreg(Surv(age, Status)~1, dist = "lognormal", data = toilet)
normalToilet = survreg(Surv(age, Status)~1, dist = "gaussian", data = toilet)
```

Toileting Weibull:

```{r}
curve(1-pweibull(x, shape = 1/0.0518, scale = exp(4.59661)), xlim= c(min(bath$age), max(bath$age)), ylab="Survival", xlab = "Age", col = "purple")
```


Toileting Exponential ?????:
```{r}
summary(expToilet)

curve(1-pexp(x, rate = 1/exp(7.2927)), xlim= c(min(toilet$age), max(toilet$age)), ylab="Survival", xlab = "Age", col = "orange")
```


Toileting Normal:
```{r}
curve(1-pnorm(x, mean = 97.8159, sd = 8), xlim= c(min(toilet$age), max(toilet$age)), ylab="Survival", xlab = "Age", col = "red")
```


Toileting Log-Normal:
```{r}
summary(logNormalToilet)

curve(1-plnorm(x, meanlog = 4.58936, sdlog = 0.0927), xlim= c(min(bath$age), max(bath$age)), ylab="Survival", xlab = "Age", col = "blue")
```


Now compare all of the parametric models to the KM curve to see which is visually the best fit:
```{r}
plot(KMtoilet, conf.int=FALSE, ylab="Survival" , xlab="Age", xlim = c(min(toilet$age), max(toilet$age)), ylim = c(0,1))

curve(1-pweibull(x, shape = 1/0.0518, scale = exp(4.59661)), col = "purple", add = TRUE)

curve(1-pexp(x, rate = 1/exp(7.2927)),  col = "orange", add = TRUE)

curve(1-pnorm(x, mean = 97.8159, sd = 8),  col = "red", add = TRUE)

curve(1-plnorm(x, meanlog = 4.58936, sdlog = 0.0927),col = "blue", add = TRUE)
```

Comparing AICs of parametric toileting models to see which parametric model fits the data best:
```{r}
#AIC = 2(P-L)

#AIC for Weibull
#summary(weibullToilet)
2*(2+1692.5)

#AIC for Exponential
#summary(expToilet)
2*(1+2595.6)

#AIC for Normal
#summary(normalToilet)
2*(2+1672.4)

#AIC for log-normal
#summary(logNormalToilet)
2*(2+1669.7)
```

AIC is lowest for the log-normal model, meaning that the log-normal model is the best fit for the toileting data.

Log-normal model for eat variable
```{r}
logNormalEat = survreg(Surv(age, Status)~1, dist = "lognormal", data = eat)

logNormalEat
```



#### Looking at the distribution for bathing and toileting
```{r}
#bathing
curve(1-plnorm(x, meanlog = 4.5439, sdlog = 0.083), col = "pink", ylab = "Survival", xlab = "Age", ylim= c(0, 1), xlim = c(min(bath$age, toilet$age, eat$age),max(bath$age, toilet$age, eat$age)))

#Toileting
curve(1-plnorm(x, meanlog = 4.58936, sdlog = 0.0927),col = "turquoise", add = TRUE)

#Eating
curve(1-plnorm(x, meanlog = 4.662521, sdlog = 0.1111955),col = "brown", add = TRUE)
```
TO DO
-do a KM curve with all of the data in one (AML) : to get a p-value for if the curves are significantly different


#### Choosing a covariate


```{r}
bath$SEX
```

```{r}
KMbathSex = survfit(Surv(age, Status)~SEX, data = bath)
KMbathSex
plot(KMbathSex, conf.int=FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col = c("blue", "red"))
```


```{r}
KMtoiletSex = survfit(Surv(age, Status)~SEX, data = toilet)
KMtoiletSex
plot(KMtoiletSex, conf.int=FALSE, ylab="Survival" , xlab="Age", xlim = c(min(toilet$age), max(toilet$age)), ylim = c(0,1), col = c("blue", "red"))
```



Race:

```{r}
bath$RACER
```

```{r}
KMbathRace = survfit(Surv(age, Status)~RACER, data = bath)
KMbathRace
plot(KMbathRace, conf.int= TRUE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col = c("blue", "red", "green"))
```

Distance away from Children (for bath dependence)
```{r}
KMbathdistance = survfit(Surv(age, Status)~children_dist, data = bath)
KMbathdistance

plot(KMbathdistance, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col = c("blue", "red", "green","orange"))
```

Distance away from Children (for toilet dependence)
```{r}
KMtoiletdistance = survfit(Surv(age, Status)~children_dist, data = toilet)
KMtoiletdistance

plot(KMtoiletdistance, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col = c("blue", "red", "green","orange"))
```

Distance away from Children (for eat dependence)
```{r}
KMeatdistance = survfit(Surv(age, Status)~children_dist, data = eat)
KMeatdistance

plot(KMeatdistance, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(eat$age), max(eat$age)), ylim = c(0,1), col = c("blue", "red", "green","orange"))
```

How often see children (for bath dependence)
```{r}
KMbathseeChildren = survfit(Surv(age, Status)~see_children, data = bath)
KMbathseeChildren

plot(KMbathseeChildren, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col = c("blue", "red", "green"))
```

How often see children (for bath dependence)
```{r}
KMbathbed = survfit(Surv(age, Status)~fam_income84, data = bath)
KMbathbed

plot(KMbathbed, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col= c("blue","red","green"))

aging$BED2
```

Questions:
1. Why are the medians NA on the distance away from children KM curves?


Regular exercise routine? (for bath dependence)
```{r}
KMbathexercise = survfit(Surv(age, Status)~REGEXE1, data = bath)
KMbathexercise

plot(KMbathexercise, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col= c("blue","red","green","orange"))

#red=YES
#blue=NO
```


Overall Health Status in 1984 for bath data
(I lumped "excellent","very good", and "good" into a GOOD category, and "fair" and "poor" into a BAD category- so I binarized it)
```{r}
KMbathhealth = survfit(Surv(age, Status)~health1984, data = bath)
KMbathhealth

plot(KMbathhealth, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col= c("blue","red"))

#red=good
#blue=bad

summary(survreg( Surv(age,Status) ~ health1984 , dist='lognormal', data=bath ))

survdiff( Surv(age,Status) ~ health1984 , data=bath )

test_stat <- 42.4+13.2

pval <- 1 - pchisq(test_stat , df=1 )

pval

(190/119)/(310/381)
```

Overall health status in 1984 for toilet data
```{r}
KMtoilethealth = survfit(Surv(age, Status)~health1984, data = toilet)
KMtoilethealth

plot(KMtoilethealth, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(toilet$age), max(toilet$age)), ylim = c(0,1), col= c("blue","red"))

#red=bad
#blue=good


```
**P-value for a Kaplan-Meier curve?? Comparing if adding a covariate matters or not**

Overall health status in 1984 for eat data
```{r}
KMeathealth = survfit(Surv(age, Status)~health1984, data = eat)
KMeathealth

plot(KMeathealth, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(eat$age), max(eat$age)), ylim = c(0,1), col= c("blue","red"))

#red=bad
#blue=good

survdiff( Surv(age,Status) ~ health1984 , data=eat )

1-pchisq(4.70+1.91,df = 1)

(56/42)/(89/103)
```



Ever had a stroke for the toilet variable (b/c the bath and eat variables had super overlapping CIs)
```{r}
KMtoiletStroke = survfit(Surv(age, Status)~stroke, data = toilet)
KMtoiletStroke

plot(KMtoiletStroke, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(toilet$age), max(toilet$age)), ylim = c(0,1), col= c("blue","red"))

#blue=no stroke
#red=yes stroke
```


```{r}
bdata = bath %>% select(Time = age, Status)
tdata = toilet %>% select(Time = age, Status)
edata = eat %>% select(Time = age, Status)
```
```{r}
BTE = rbind(bdata, tdata, edata)
BTE = BTE %>% mutate(Group = c(rep("bath", 5204), rep("toilet", 5616), rep("eat", 5901)))
```

```{r}
survreg( Surv(Time,Status) ~ Group , dist='lognormal', data=BTE )
```

EDUCATION 
*remember to select
```{r}
bath$EDUCAT= ifelse(bath$EDUC %in% c("09", "10", "11", "12"), 2, ifelse(bath$EDUC %in% c("13", "14", "15", "16", "17", "18"), 3, ifelse(bath$EDUC == "19", NA, 1)))
```


```{r}
newBath = bath %>% filter(EDUCAT %in% c(1, 2, 3))
```


```{r}
km = survfit(Surv(age, Status)~EDUCAT, data = newBath)
km

plot(km, conf.int= FALSE, ylab="Survival" , xlab="Age", xlim = c(min(bath$age), max(bath$age)), ylim = c(0,1), col= c("blue","red","green"))

```



#==================================

Looking into the age distribution of bathroom dependence

```{r}
#Creating a new variable called BATHDEPYR to show when they began dependent
bathData$BATHDEPYR <- ifelse(bathData$BATH3R == "2", 1988, 1990)
```


```{r}
#Showing which are censored data
bathData$status <- ifelse(bathData$BATH4R == "1", 0, 1)
```


```{r}
#age to bath dependence
bathData$AGE <- (bathData$BATHDEPYR - as.integer(as.character(bathData$BIRTHYR)))

#If the person did not want to answer their birthday year would have been 9999

bathData <- filter(bathData, bathData$AGE > 0) 
```



```{r}
bathData %>%
  select(TLKCHL2) %>%
  table()
```



```{r}
#The following is a Kaplan Meier curve for the time a person becomes dependent with taking a shower
library(survival)
bathKM= survfit(Surv(bathData$AGE, bathData$status) ~ RACE , data=bathData )
plot(bathKM, conf.int=FALSE, ylab="Survival Proportion" , xlab="Years", xlim = c(min(bathData$AGE), max(bathData$AGE)), ylim = c(0,1))

```


```{r}
bathKM
```


The median age to have bathroom dependence is 88 years
```{r}
#The median age of having bathroom dependence 
bathKM
```


The mean age of having bathroom dependence is 87.94954 years
```{r}
#Using Victor's code to find the mean age of having bathroom dependence

AUCKM = function(survobj,duration)
{
base=c(0,summary(survobj)$time,max(duration))
heights=c(1,summary(survobj)$surv)
new=c()
for(i in 1:length(heights)) { new=c(new,(base[i+1]-base[i])*heights[i]) }
c(sum(new))
}

AUCKM(bathKM, bathData$AGE)
```






dependence:
-treat "dontknow/refuse" as missing
-if they aren't dependent in 1986: keep
-if they are dependent in 1986: use only if necessary (if not enough data)

PROCESS
-check when the 1s change to 2s
-record the year that that happens
-subtract the birthyear (gives us age when they became dependent) and make status variable ==1

-if it never becomes 2
-take last survey year
-subtract birthyear and make status ==0

-we will have age/status pair for each dependency kind

GOAL:
-do procress for bathing


Some project analysis ideas are:
-create survival curves for specific covariates (get the CI, mean, and median) using KM curves
- From the KM curves that look the most interesting, choose the parametric model that fits it best (Cox-Snell(graphically check), Likelihood Ratio Test, AIC)
- Compare variables to each other (Log Rank Test and Hazard Ratio )


Some covariate could potentilly be time-varying
-Recurrent Events (look into this)
